Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars II

### 2023-07-24 project/overleaf/paper2/mblithium.tex
WIP paper II

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[normalem]{ulem}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{longtable}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

\section{Data} \label{sec_method}
The datasets provided by the Gaia mission in GEDR3 and GDR3, as well as those published by GES in DR5.0 form the main references used in this work. In general, we focus on the observations made on OCs, and particularly on those analysed in the \cite{Bragaglia2022} and \cite{Randich2022} papers. On the basis of the results presented in these papers, a set of 65 OCs (see Table \ref{tab:oc_list}) are selected as the reference dataset for our work.

@@

  	\label{tab:oc_list}

@@

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            8510914\_1157003 & 5843 & 4.42 & -0.03 & 1.8 & 0.08 & 3.981\\ 
            8510991\_1146169 & 5797 & 4.4 & -0.01 & 1.65 & 0.08 & 3.981\\ 
            8512177\_1144050 & 5818 & 4.46 & 0 & 1.99 & 0.06 & 3.981\\ 
            8513941\_1200571 & 5792 & 4.41 & -0.02 & 1.49 & 0.13 & 3.981\\ 
            8515559\_1148381 & 5794 & 4.43 & -0.04 & 1.56 & 0.1 & 3.981\\ 
            8520350\_1147480 & 5875 & 4.47 & -0.03 & 1.21 & 0.24 & 3.981\\ 
            \hline
	\end{tabular}
 	\caption{M64 - Filter setup: $\teff$ low=5789.184, $\teff$ top=5889.6435, $\gsurf$ low=4.3945, $\gsurf$ top=4.4939, $\feh$ low=-0.05, $\feh$ top=0.05, Age low=3.977, Age top=3.985}
	\label{tab:oc_m64}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            12392824-6041134 & 5831 & 4.54 & 0.01 & 2.66 & 0.08 & 1.862\\ 
            12401318-6037507 & 5810 & 4.54 & 0 & 2.35 & 0.12 & 1.862\\ 
            12402039-6042428 & 5824 & 4.48 & 0.02 & 2.05 & 0 & 1.862\\ 
            \hline
	\end{tabular}
 	\caption{Trumpler20 - Filter setup: $\teff$ low=5731.2886, $\teff$ top=5831.2886, $\gsurf$ low=4.4518, $\gsurf$ top=4.5518, $\feh$ low=-0.05, $\feh$ top=0.05, Age low=1.8601, Age top=1.8639}
	\label{tab:oc_trumpler20}

As a result of this first filtering, the number of selected records is reduced to 43 299. On this subset of data we perform another filtering, this time aimed at selecting those that offer information on the following attributes:
\begin{itemize}
    \item Metallicity
    \item Li abundance
    \item GES field identifier
    \item Effective temperature
    \item Surface gravity
\end{itemize}

This second filtering reduces the population of stars to be considered to a total of 21 125 records. Finally, a third filtering is carried out to select those stars belonging to the OCs listed in Table \ref{tab:oc_list}, giving us a final study population of 20 913 stars distributed in 65 OCs.\par

In our case we extend the selection criteria by adding additional constraints: the surface gravity and age of the star. These two new parameters represent a fine-tuning of the search process for sibling candidates to our Sun. We use the age of the OCs reported in Table \ref{tab:oc_list} and the data obtained from the set of simulations performed using Modules for Experiments in Stellar Astrophysics stellar evolution code \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}. Our models extend the investigations documented in \cite{Navarro2020} on the influence of rotation and magnetic braking on the star Li depletion during the pre-main sequence (PMS) and the main sequence (MS). The data obtained from the simulations are confronted by comparing different stellar parameters obtained by Gaia and GES. This allows us to analyse, validate and refine our models against a very large set of candidates. The end result of this process is to obtain a subset of candidate sibling stars of our Sun that help us to support (or refute) our line of investigation of how the Li abundance behaves throughout the evolution of solar-type stars.\par

From the approximately 21 000 identified stars, we proceed to select those with metallicity ($\feh$), effective temperature ($\teff$), and surface gravity ($\gsurf$) within the following limits:
\begin{itemize}
    \item $0.05 \, \dex <= \feh <= 0.05 \, \dex$
    \item $50.0 \, \Kelvin <= \teff <= 50.0 \, \Kelvin$
    \item $0.05 \, \dex <= \gsurf <= 0.05 \, \dex$
\end{itemize}

Tables \ref{tab:oc_m64} and \ref{tab:oc_trumpler20} show the subset of components of OC M67 and Trumpler 20 selected during the simulations for models initialised with $\oomegac$ = 0.14. Each of the tables lists the GES identifier of the star and the $\teff$, $\gsurf$, $\feh$, A(Li) and estimated age ranges used in the simulation.

In the remainder of this paper, we describe a semi-empirical approach for the calculation of the AML as a result of the braking caused by a magnetic field of varying strength. The magnetic field evolves according to the $\teff$ and angular velocity ($\Omega$) of the star. This has been implemented as an extension to MESA.

\subsection{Modelling magnetic intensity} \label{mod_mbi}
The variable magnetic field strength modelling approach is supported by the papers \cite{Cranmer2011} and \cite{Gallet2013}. We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of magnetic field intensity.\par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is $\Omega$. Assuming that the magnetic field is generated by a dynamo, its strength is proportional to some power of $\Omega$,
\begin{ceqn}
\begin{equation}
    f_*B_* \propto \Omega_*^b \label{eq:mf_strenght}
\end{equation}
\end{ceqn}
where $b$ is the dynamo exponent, $B_*$ is the strength of the magnetic filed, and $f_*$ is the filling factor, i.e. the fraction of the star surface which is magnetized \citep{Gallet2013}.

On the other hand, $B_*$ is proportional to the equipartition magnetic field strength,
\begin{ceqn}
\begin{equation}
    B_* \approx 1.13 \Beq \label{eq:mf_bstar}
\end{equation}
\end{ceqn}
where $\Beq$ is defined as 
\begin{ceqn}
\begin{equation}
    \Beq = \sqrt{8\upi P_*} = \sqrt{\frac{8\upi\rho_* \boltzmann \teff}{\mu\massH}}\label{eq:mf_beq}    
\end{equation}
\end{ceqn}
where $\rho_*$ is the photospheric density, $\boltzmann$ is Boltzmann's constant, $\massH$ is the mass of a hydrogen atom, and $\mu$ the mean atomic weight \citep{Cranmer2011}.

\cite{Cranmer2011} found that the magnetic field strength is mainly influenced by $f_*$ rather than by the rotation period of the star. That is, variations in the star's angular velocity do not significantly alter its magnetic field strength. In addition, $f_*$ has a strong dependence on the Rossby number ($\rossby$). In order to calculate $\rossby$ of a given star, it is necessary to know the turn over time ($\turnover$),
\begin{ceqn}
\begin{equation}
    \rossby = \frac{\prot}{\turnover} \label{eq:mf_rossby}
\end{equation}
\begin{equation}
    \turnover = 314.24\;exp\left[-\Bigg(\frac{\teff}{1952.5}\Bigg)-\Bigg(\frac{\teff}{6250}\Bigg)^{18} \;\right]+0.002 \label{eq:mf_turnover}
\end{equation}
\end{ceqn}
\cite{Cranmer2011} Crammer provides two different settings for $f_*$ which define respectively the upper and lower limits for the mean surface magnetic field strength ($f_*B_*$),

Finally, we are able to calculate the filling factor for a star by applying the adjustment made by \cite{Gallet2013},
\begin{ceqn}
\begin{equation}
     \fstar = \frac{0.55}{(1+(x/0.16)^{2.3})^{1.22}} \label{eq:fstart}
\end{equation}
\end{ceqn}

\subsection{Modelling magnetic braking} \label{mod_mb}
Once we have characterised how the magnetic field strength evolves, we need to model how the loss of angular momentum occurs.  In \cite{Navarro2020} we use a model applicable to magnetic fields of non-variable intensity and of few Gauss ($1.0\,\Gauss \leq B \leq 10.0\,\Gauss$). As shown in Figure \ref{fig:mag_field_limits_var_vel_g3}, the simulations yield magnetic field values that show maximum intensities of around $1.0K\,\Gauss$, making the use of the model employed previously non viable.

\subsection{Mesh Models} \label{sec_mesh}
In the current work we have replaced two pre-set parameters (alpha and intensity) for the different simulations of our model. Not only have we eliminated the need to assign a value to them, but we allow their values to vary as the simulations evolve over time. The single free parameter of our implementation is $\Omega$. The numerical simulations traced the rotational history and $A(\isotope[7]{Li})$ of a $1\, \msun$ star for a variety of initial values for $\Omega$ (see also Table \ref{tab:phy_mesa}).\par

The distribution is performed in two iterations. In the first iteration, each cell $k$ is assigned the maximum amount of $\Dot{J}_{k}$ that it can accommodate in the time step established by the simulation. The amounts of $\Dot{J}$ not accommodated by the cells are accumulated as residual angular moment ($\jres$). In a second pass, the aggregated residual $\jres$ value collected in the previous iteration is redistributed among those cells that can accommodate more torque. In this way we manage to conserve the angular momentum.\par

\begin{figure}
	\includegraphics[trim = 25mm 10mm 15mm 10mm, clip, width=\columnwidth]{figures/paper1/li_var_vel_0_0g.eps}
    \caption{The evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H}, as a function of time for several 1 $\msun$ models. The solid black line represents the reference model according to \citet{Choi2016}. The rest of lines are models which include PMS rotation with $\oomegac$ between 0.0084 and 0.0336, respectively. The purple star and square are surface Li abundances for the present-day Sun \citep{Asplund2009} and the average for the Pleiades cluster \citep{Sestito2005} respectively. The dashed vertical line makes reference to the ZAMS.}
    \label{fig:li_var_vel_0g}
\end{figure}

\begin{figure}
	\includegraphics[trim = 25mm 10mm 15mm 10mm clip,width=\columnwidth]{figures/paper1/li_var_vel_0_0g_z1.eps}
    \caption {Similar to Figure \ref{fig:li_var_vel_0g} but zooming in on the ZAMS. The models with a higher initial rotational velocity reach already the ZAMS with a lower amount of Li measure on the star surface.}
    \label{fig:li_var_vel_0g_z1}
\end{figure}

\begin{figure}
	\includegraphics[trim = 10mm 10mm 15mm 10mm clip,width=\columnwidth]{figures/paper1/rot_vel_var_vel_0_0g.eps}
    \caption{The evolution of angular velocity at surface (solid line) and at the bottom (dotted line) limit of the uppermost convective zone, as a function of time for several 1 $\msun$ models. The models includes PMS rotation with $\oomegac$ values between 0.0084 and 0.0336. The purple star is the surface angular velocity for the present-day Sun \citep{Gill2012}. The dashed vertical line makes reference to the ZAMS.}
    \label{fig:rot_vel_0g}
\end{figure}

\begin{figure}
	\includegraphics[trim = 10mm 10mm 15mm 10mm, clip,width=\columnwidth]{figures/paper1/hr_var_vel_0_0g_z1.eps}
    \caption{An example solar 1$\msun$ grid of stellar evolutionary track from PMS till TAMS covering a wide range of angular velocities. The rotation is activated in the models in the PMS and those models reach before the ZAMS and at a lower $\teff$ than the non-rotating one (solid black line). The luminosity is expressed in terms of $\lsun$.}
    \label{fig:hr_var_vel_0g}
\end{figure}

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix
\clearpage
\section{Mesh models visualization}
The following figures comprise a series of grids as a function of time and for several 1 $\msun$ models in which on the one hand, the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H} for both variable magnetic field intensities and angular velocities and on the other hand, the evolution of surface rotational velocity are shown.

\begin{figure*}
    \centering
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[clip,width=\textwidth]{figures/paper2/li_var_vel_var_g_1.pdf}
    \label{fig:subim11}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[clip,width=\textwidth]{figures/paper2/li_var_vel_var_g_3.pdf}
    \label{fig:subim12}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[trim = 25mm 10mm 15mm 10mm, clip,width=\textwidth]{figures/paper1/li_var_vel_0_0g.eps}
    \label{fig:subim13}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[width=\textwidth]{figures/blank.eps}
    \label{fig:subim14}
    \end{subfigure}

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[normalem]{ulem}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{longtable}
\usepackage{threeparttable}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

\section{Data} \label{sec_method}
The data sets provided by the Gaia mission in GEDR3 and GDR3, as well as those published by GES in DR5.0 form the main references used in this work. In general, we focus on the observations made on OCs, and particularly on those analysed in the \cite{Bragaglia2022} and \cite{Randich2022} papers. On the basis of the results presented in these papers, a set of 64 OCs (see Table \ref{tab:oc_full_list}) are selected as the reference dataset for our work.

\begin{table*}
	\centering
	\begin{tabular}{|l l l l l || c c c c c | c c c c c|} 
		\hline
             & & & & & & & $\oomegac$ & & & & & $\oomegac$ & & \\
		Open Cluster & GES\_FLD & Age (Gyr) & [Fe/H] & $N_*$ & 0.095 & 0.10 & 0.105 & 0.11 & 0.115 & 0.12 & 0.125 & 0.13 & 0.14 & 0.1425\\
		\hline
            Berkeley 21 & Br21 & 2.138 & -0.21 & 744 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1\\
            Berkeley 39 & Br39 & 5.623 & -0.14 & 899 & 1 & 1 & 1 & 1 & 1 & 1 & 2 & 2 & 2 & 2\\
            IC 2602 & IC2602 & 0.036 & -0.06 & 1836 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0\\
            IC 4665 & IC4665 & 0.033 & 0.01 & 567 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0\\
            Messier 67 & M67 & 3.981 & -0.02 & 131 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6\\
            NGC 2141 & NGC2141 & 1.862 & -0.04 & 853 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2355 & NGC2355 & 1 & -0.13 & 208 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2420 & NGC2420 & 1.698 & -0.15 & 562 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2425 & NGC2425 & 2.399 & -0.13 & 528 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2451 & NGC2451 & 0.035 & -0.08 & 1656 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 2 & 1\\
            NGC 2516 & NGC2516 & 0.24 & -0.04 & 759 & 0 & 0 & 0 & 1 & 1 & 1 & 3 & 4 & 4 & 3\\
            NGC 3532 & NGC3532 & 0.398 & -0.01 & 1145 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 1\\
            NGC 6005 & NGC6005 & 1.259 & 0.22 & 560 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1\\
            NGC 6259 & NGC6259 & 0.269 & 0.18 & 494 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
            NGC 6281 & NGC6281 & 0.513 & -0.04 & 320 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            NGC 6405 & NGC6405 & 0.035 & -0.02 & 701 & 1 & 1 & 1 & 0 & 1 & 3 & 2 & 0 & 0 & 0\\
            NGC 6633 & NGC6633 & 0.692 & -0.03 & 1662 & 2 & 2 & 2 & 1 & 1 & 0 & 0 & 0 & 1 & 1\\
            NGC 6709 & NGC6709 & 0.191 & -0.02 & 730 & 2 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            Trumpler 20 & Trumpler20 & 1.862 & 0.13 & 1213 & 3 & 3 & 3 & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
            \hline
	\end{tabular}
 	\caption{List of selected OCs. For each OC, name, GES denomination, estimated age, metallicity and number of components are listed. The number of OC components selected by the different simulations are shown separately. This information is split into two groups, each corresponding to the set of simulations used in the figures shown in the following sections.}
  	\label{tab:oc_reduced_list}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            8510914\_1157003 & 5843 & 4.42 & -0.03 & 1.8 & 0.08 & 3.981\\ 
            8510991\_1146169 & 5797 & 4.4 & -0.01 & 1.65 & 0.08 & 3.981\\ 
            8512177\_1144050 & 5818 & 4.46 & 0 & 1.99 & 0.06 & 3.981\\ 
            8513941\_1200571 & 5792 & 4.41 & -0.02 & 1.49 & 0.13 & 3.981\\ 
            8515559\_1148381 & 5794 & 4.43 & -0.04 & 1.56 & 0.1 & 3.981\\ 
            8520350\_1147480 & 5875 & 4.47 & -0.03 & 1.21 & 0.24 & 3.981\\ 
            \hline
	\end{tabular}
 	\caption{M67 - Filter setup: $\teff$(K) [5787.4892, 5887.6823], $\gsurf$(dex) [4.3979, 4.4977], $\feh$(dex) [-0.05, 0.05], Age(Gyr) [3.977, 3.985]}
	\label{tab:oc_m67}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            7571111-6048156 & 5084 & 4.54 & 0 & 2.29 & 0.08 & 0.24\\ 
            7595031-6044149 & 5108 & 4.52 & -0.05 & 1.92 & 0.08 & 0.24\\ 
            7595280-6032498 & 5093 & 4.49 & -0.03 & 2.29 & 0.08 & 0.24\\ 
            \hline
	\end{tabular}
 	\caption{NGC2516 - Filter setup: $\teff$(K) [5010.0324, 5111.0095], $\gsurf$(dex) [4.4405, 4.5405], $\feh$(dex) [-0.05, 0.05], Age(Gyr) [0.23964, 0.24036]}
	\label{tab:oc_ngc2516}

As a result of this first filtering, the number of selected records is reduced to 43 299. On this subset of data we perform another filtering, this time aimed at selecting those that offer not null information on the following attributes:
\begin{itemize}
    \item Metallicity
    \item Li abundance
    \item GES field identifier
    \item Effective temperature
    \item Surface gravity
    \item Cluster membership probability (>= 0.95)
\end{itemize}

This second filtering reduces the population of stars to be considered to a total of 5 895 records which are distributed among the 58 OCs listed in Table \ref{tab:oc_full_list}.\par

In our case we extend the selection criteria by adding additional constraints: the surface gravity and age of the star. These two new parameters represent a fine-tuning of the search process for sibling candidates to our Sun. We use the age of the OCs reported in Table \ref{tab:oc_reduced_list} and the data obtained from the set of simulations performed using Modules for Experiments in Stellar Astrophysics stellar evolution code \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}. Our models extend the investigations documented in \cite{Navarro2020} on the influence of rotation and magnetic braking on the star Li depletion during the pre-main sequence (PMS) and the main sequence (MS). The data obtained from the simulations are confronted by comparing different stellar parameters obtained by Gaia and GES. This allows us to analyse, validate and refine our models against a very large set of candidates. The end result of this process is to obtain a subset of candidate sibling stars of our Sun that help us to support (or refute) our line of investigation of how the Li abundance behaves throughout the evolution of solar-type stars.\par

From the approximately 5 900 identified stars, we proceed to select those with metallicity ($\feh$), effective temperature ($\teff$), and surface gravity ($\gsurf$) within the following limits:
\begin{itemize}
    \item $0.05 \, \dex <= \feh <= 0.05 \, \dex$
    \item $50.0 \, \Kelvin <= \teff <= 50.0 \, \Kelvin$
    \item $0.05 \, \dex <= \gsurf <= 0.05 \, \dex$
\end{itemize}

The method of selection is as follows: Based on the values for $\teff$, $\gsurf$ and $\feh$ given by our simulations, we proceed to select in the different OCs those stars that have values within the established ranges. That is, for the estimated age of a particular cluster, we calculate an interval of $\pm$0.1 for young OCs (age < 1Gyr) and $\pm$0.15 for old OCs (age >= 1Gyr) \citep[see][and reference values therein]{Cantat-Gaudin2020}. This age interval is used to select in our simulations the time step that is closest to the lower and upper age limits. Once these two time steps are found, the associated limit values for $\teff$, $\gsurf$ and $\feh$ are gathered. With these value ranges we proceed then to select in each of the OCs those stars that have values for these same parameters within them. In this way, we finally obtain the candidate stars for relevant for being compared with the simulation results.\par

Tables \ref{tab:oc_m67} and \ref{tab:oc_ngc2516} summarize the subset of components of OC M67 and NGC2516 selected during the simulations for models initialised with $\oomegac$ = 0.14. Each of the tables lists the GES identifier of the star and the $\teff$, $\gsurf$, $\feh$, A(Li) and estimated age ranges used in the simulation.

In \cite{Gossage2021}, the authors also approach the problem of AML from a similar starting point to ours, but with some notable differences. In the Table \ref{tab:gassage_vs_navarro} we compare the most notable similarities and differences between the two studies. In both works the models consider rotation not only in the MS but also in the PMS, with the difference that in Gossage's work they take into account disk-locking. This aspect is not so important in our simulations since in our case we have focused on the A(Li) profiles. In contrast, Gossage's work focuses mainly on modelling rotation profiles of stars, which are significantly affected by disk-locking. Similarly, the effects of magnetic braking on angular momentum loss are also considered. In our case for stars with $\mstar$=1$\msun$ (solar-twins) and in their with 0.1$\msun$$\le$$\mstar$$\le$1.3$\msun$. Both models also use MLT but in our case the MLT alpha value is not pre-fixed but varies throughout the simulation. These two works complement each other as they are based on different formalisms. On the one hand, our work is based on the proposal of \cite{Gallet2013} whose most relevant parameters are Ro and the presence of a magnetic field of varying strength (B), and on the other hand Gossage presents two formalisms: \cite{Matt2015} and \cite{Garraffo2018}. The former is mainly characterised by $\rossby$ and the $\mstar$, while the latter by $\rossby$ and the complexity of the magnetic field ($\magfc$). As far as OCs observations are concerned, in this work we have had access to a larger number of them, and covered a wider range of ages [0.001-6.761] Gyr. Finally, both papers simulate their respective models in MESA.\par

\begin{table}
	\centering
	\caption{Comparison with \cite{Gossage2021} work.}
	\label{tab:gassage_vs_navarro}
        \begin{threeparttable}
	\begin{tabular}{lll} 
		\hline
		Parameter & This work & \cite{Gossage2021}\\
		\hline

		Rotation & At PMS \& MS & At PMS \& MS\\
  		Solar Mass & $\mstar$=1$\msun$ & 0.1$\msun$$\le$$\mstar$$\le$1.3$\msun$ \\
		Convection & MLT variable $\alpha_{{\rm MLT}}$ & MLT fixed $\alpha_{{\rm MLT}}$\\
		Magnetic Braking & \cite{Gallet2013} & \cite{Matt2015}, \\ Formalism & & \cite{Garraffo2018}\\
            Magnetic Braking & $\rossbystar$ \& $\bstar$ & $\rossbystar$ \& $\mstar$, \\ Parameters & & $\rossbystar$ \& $\magfc$\tnote{(1)}\\
		Magnetic Field & B(G) variable & Not used\\ Strength & & \\
            Disk locking & Not present & Present\\
            Number of OC's & 64 & 11 \\
            Source of data & GDR3.0 GES DR5.0 & Several works \\
            MESA version & r10398 & r11701\\
		\hline
	\end{tabular}
        \end{threeparttable}
        \begin{tablenotes}\footnotesize
        \item (1) Magnetic field complexity
        \end{tablenotes}
\end{table}

In the remainder of this paper, we describe a semi-empirical approach for the calculation of the AML as a result of the braking caused by a magnetic field of varying strength. The magnetic field evolves according to the $\teff$ and angular velocity ($\Omega$) of the star. This has been implemented as an extension to MESA.\par

\subsection{Modelling magnetic intensity} \label{mod_mbi}
In our models we make the case that the magnetic field strength should vary over the lifetime of the star rather than remain fixed throughout its evolution. For this we rely on a formalism \citep{Gallet2013} that allows us to calculate a variable magnetic field strength. This model adopts a dynamo-type prescription that allows a relationship to be established between the $\Omega$ , $\teff$, and the strength of the magnetic field ($B$), as well as with the mass-loss ($\mwind$) rate caused by the stellar wind. We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of magnetic field intensity. \\par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is $\Omega$. Assuming that the magnetic field is generated by a dynamo, its strength is proportional to some power of $\Omega$,
\begin{ceqn}
\begin{equation}
    f_*B_* \propto \Omega_*^b \label{eq:mf_strenght}
\end{equation}
\end{ceqn}
where $f_*$ is the filling factor, i.e. the fraction of the star surface which is magnetized, $B_*$ is the strength of the magnetic filed, $\Omega_*$ is the star angular velocity at stellar surface, and $b$ is the dynamo exponent \citep{Gallet2013}. The multiplication of $f_*$ by $B_*$ will allow us to obtain the magnetic field strength. \par

According to this approach, the magnetic field strength will vary proportionally to the effective temperature ($\teff$) of the star as follows: 
\begin{ceqn}
\begin{equation}
    B_* \approx 1.13 \Beq \label{eq:mf_bstar}
\end{equation}
\end{ceqn}
$B_*$ is proportional to the equipartition magnetic field strength, being $\Beq$ is defined as 
\begin{ceqn}
\begin{equation}
    \Beq = \sqrt{8\upi P_*} = \sqrt{\frac{8\upi\rho_* \boltzmann \teff}{\mu\massH}}\label{eq:mf_beq}    
\end{equation}
\end{ceqn}
where $P_*$ is the photospheric gas pressure, $\rho_*$ is the photospheric density, $\boltzmann$ is Boltzmann's constant, $\massH$ is the mass of a hydrogen atom, and $\mu$ the mean atomic weight \citep{Cranmer2011}.

On the other hand, \cite{Cranmer2011} found that the magnetic field strength is mainly influenced by $f_*$ rather than by the rotation period of the star. That is, variations in the star's angular velocity do not significantly alter its magnetic field strength. In addition, $f_*$ has a strong dependence on the Rossby number ($\rossby$). In order to calculate $\rossby$ of a given star, it is necessary to know the turn over time ($\turnover$),
\begin{ceqn}
\begin{equation}
    \rossby = \frac{\prot}{\turnover} \label{eq:mf_rossby}
\end{equation}
\begin{equation}
    \turnover = 314.24\;exp\left[-\Bigg(\frac{\teff}{1952.5}\Bigg)-\Bigg(\frac{\teff}{6250}\Bigg)^{18} \;\right]+0.002 \label{eq:mf_turnover}
\end{equation}
\end{ceqn}
\cite{Cranmer2011} Crammer provides two different settings for $f_*$ which define respectively the upper and lower limits for the mean surface magnetic field strength ($f_*B_*$),

Finally, we are able to calculate the filling factor for a star by applying the adjustment made by \cite{Gallet2013} which more closely reproduces the average filling factor of the present Sun ($\fsun$ = 0.001-0.01),
\begin{ceqn}
\begin{equation}
     \fstar = \frac{0.55}{(1+(x/0.16)^{2.3})^{1.22}} \label{eq:fstart}
\end{equation}
\end{ceqn}

\subsection{Modelling magnetic braking} \label{mod_mb}
Once we have characterised how the magnetic field strength evolves, we need to model how the loss of angular momentum occurs.  In \cite{Navarro2020} we use a model applicable to magnetic fields of non-variable intensity and of few Gauss ($1.0\,\Gauss \leq B \leq 10.0\,\Gauss$). As shown in Figure \ref{fig:mag_field_limits_var_vel_g3}, the simulations yield magnetic field values that show maximum intensities of around $1.0K\,\Gauss$, making the use of the model employed previously non viable. \par

\subsection{Grid Models} \label{sec_grid}
In the current work we have replaced two pre-set parameters (alpha and intensity) for the different simulations of our model. Not only have we eliminated the need to assign a value to them, but we allow their values to vary as the simulations evolve over time. The single free parameter of our implementation is $\Omega$. The numerical simulations traced the rotational history and $A(\isotope[7]{Li})$ of a $1\, \msun$ star for a variety of initial values for $\Omega$ (see also Table \ref{tab:phy_mesa}).\par

\begin{figure}
	\includegraphics[trim = 25mm 10mm 15mm 10mm, clip, width=\columnwidth]{figures/paper1/li_var_vel_0_0g.pdf}
    \caption{The evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H}, as a function of time for several 1 $\msun$ models. The solid black line represents the reference model according to \citet{Choi2016}. The rest of lines are models which include PMS rotation with $\oomegac$ between 0.0084 and 0.0336, respectively. The purple star and square are surface Li abundances for the present-day Sun \citep{Asplund2009} and the average for the Pleiades cluster \citep{Sestito2005} respectively. The dashed vertical line makes reference to the ZAMS.}
    \label{fig:li_var_vel_0g}
\end{figure}

\begin{figure}
	\includegraphics[trim = 25mm 10mm 15mm 10mm clip,width=\columnwidth]{figures/paper1/li_var_vel_0_0g_z1.pdf}
    \caption {Similar to Figure \ref{fig:li_var_vel_0g} but zooming in on the ZAMS. The models with a higher initial rotational velocity reach already the ZAMS with a lower amount of Li measure on the star surface.}
    \label{fig:li_var_vel_0g_z1}
\end{figure}

\begin{figure}
	\includegraphics[trim = 10mm 10mm 15mm 10mm clip,width=\columnwidth]{figures/paper1/rot_vel_var_vel_0_0g.pdf}
    \caption{The evolution of angular velocity at surface (solid line) and at the bottom (dotted line) limit of the uppermost convective zone, as a function of time for several 1 $\msun$ models. The models includes PMS rotation with $\oomegac$ values between 0.0084 and 0.0336. The purple star is the surface angular velocity for the present-day Sun \citep{Gill2012}. The dashed vertical line makes reference to the ZAMS.}
    \label{fig:rot_vel_0g}
\end{figure}

\begin{figure}
	\includegraphics[trim = 10mm 10mm 15mm 10mm, clip,width=\columnwidth]{figures/paper1/hr_var_vel_0_0g_z1.pdf}
    \caption{An example solar 1$\msun$ grid of stellar evolutionary track from PMS till TAMS covering a wide range of angular velocities. The rotation is activated in the models in the PMS and those models reach before the ZAMS and at a lower $\teff$ than the non-rotating one (solid black line). The luminosity is expressed in terms of $\lsun$.}
    \label{fig:hr_var_vel_0g}
\end{figure}

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Full list of OC's}

@@

  	\label{tab:oc_full_list}

@@

\clearpage
\section{Grid models visualization}
The following figures comprise a series of grids as a function of time and for several 1 $\msun$ models in which on the one hand, the evolution of surface \isotope[7]{Li} abundance relative to \isotope[1]{H} for both variable magnetic field intensities and angular velocities and on the other hand, the evolution of surface rotational velocity are shown.

\begin{figure*}
    \centering
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[clip,width=\textwidth]{figures/paper2/li_var_vel_var_g_1.pdf}
    \label{fig:subim11}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[clip,width=\textwidth]{figures/paper2/li_var_vel_var_g_3.pdf}
    \label{fig:subim12}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[trim = 25mm 10mm 15mm 10mm, clip,width=\textwidth]{figures/paper1/li_var_vel_0_0g.pdf}
    \label{fig:subim13}
    \end{subfigure}
    \begin{subfigure}[h]{0.47\textwidth}
    \includegraphics[width=\textwidth]{figures/blank.eps}
    \label{fig:subim14}
    \end{subfigure}
